---
title: "data_analysis"
output: html_document
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tidycensus)
library(tigris)
library(sf)
```


```{r}
clergy_contribs_2016 <- read_csv("clergy_contribs_2016.csv")
clergy_contribs_2018 <- read_csv("clergy_contribs_2018.csv")
clergy_contribs_2020 <- read_csv("clergy_contribs_2020.csv")
clergy_contribs_2022 <- read_csv("clergy_contribs_2022.csv")
clergy_contribs_2024 <- read_csv("clergy_contribs_2024.csv")

clergy_contribs_combined <- bind_rows(list(clergy_contribs_2016, 
                                           clergy_contribs_2018, 
                                           clergy_contribs_2020, 
                                           clergy_contribs_2022, 
                                           clergy_contribs_2024)) |> 
  mutate(zip5 = str_sub(contributor_zip, start=1L, end=5L))

clergy_totals_by_cycle <- clergy_contribs_combined |>
  group_by(contributor_name, two_year_transaction_period, contributor_employer, zip5, committee_name) |>
  summarize(
    total_amount = sum(contribution_receipt_amount),
    num_contributions = n(),
  )

clergy_wide <- clergy_totals_by_cycle |>
  select(contributor_name, contributor_employer, two_year_transaction_period, total_amount, zip5, committee_name) |>
  pivot_wider(
    names_from = two_year_transaction_period,
    values_from = total_amount,
    values_fill = 0) |> 
  rename(denomination = contributor_employer) |> 
  mutate(
    denomination = case_when(
        str_detect(denomination, "BAPTIST") ~ "Baptist",
        str_detect(denomination, "CATHOLIC") ~ "Catholic",
        str_detect(denomination, "METHODIST") ~ "Methodist",
        str_detect(denomination, "PRESBY") ~ "Presbyterian",
        str_detect(denomination, "LUTHERAN") ~ "Lutheran",
        str_detect(denomination, "EPISCOPAL") ~ "Episcopal",
        str_detect(denomination, "AME") ~ "AME",
        str_detect(denomination, "ORTHODOX") ~ "Orthodox",
        str_detect(denomination, "CHURCH OF CHRIST") ~ "Church of Christ",
        str_detect(denomination, "ADVENTIST") ~ "Seventh-Day Adventist",
        str_detect(denomination, "SYNAGOGUE") ~ "Jewish",
        str_detect(denomination, "BUDDHIST") ~ "Buddhist",
        str_detect(denomination, "HINDU") ~ "Hindu",
        str_detect(denomination, "SIKH") ~ "Sikh",
        str_detect(denomination, "RETIRED") ~ "Retired",
        str_detect(denomination, "SELF") ~ "Self Employed",
        str_detect(denomination, "ARCHDIOCESE") ~ "Catholic",
        str_detect(denomination, "CONFERENCE") ~ "Seventh-Day Adventist",
        .default = NA
      )
  ) |> 
  mutate(
    affiliation = case_when(
      str_detect(committee_name, "ACTBLUE") ~ "DEMOCRAT",
      str_detect(committee_name, "HILLARY") ~ "DEMOCRAT",
      str_detect(committee_name, "KAMALA") ~ "DEMOCRAT",
      str_detect(committee_name, "BIDEN") ~ "DEMOCRAT",
      str_detect(committee_name, "FETTERMAN") ~ "DEMOCRAT",
      str_detect(committee_name, "DSCC") ~ "DEMOCRAT",
      str_detect(committee_name, "DEMOCRATIC") ~ "DEMOCRAT",
      str_detect(committee_name, "STOP REPUBLICAN") ~ "DEMOCRAT",
      str_detect(committee_name, "STOP") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "DCCC") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "ALSOBROOKS") ~ "DEMOCRAT",
      str_detect(committee_name, "BERNIE") ~ "DEMOCRAT",
      str_detect(committee_name, "RASKINS") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "REPUBLICAN & CONVENTION") ~ "REPUBLICAN",
      str_detect(committee_name, "MAKE AMERICA GREAT AGAIN") ~ "REPUBLICAN",
      str_detect(committee_name, "TRUMP") ~ "REPUBLICAN",
      str_detect(committee_name, "WINRED") ~ "REPUBLICAN",
      str_detect(committee_name, "RAND") ~ "REPUBLICAN",
      str_detect(committee_name, "REPUBLICAN NATIONAL COMMITTEE") ~ "REPUBLICAN",
      str_detect(committee_name, "NEVER SURRENDER") ~ "REPUBLICAN",
      .default = NA
    )
)
  

total_denomination_contribs <- clergy_wide |> 
  group_by(denomination) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))

total_party_affiliation <- clergy_wide |> 
  group_by(affiliation) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))

write_csv(total_denomination_contribs, "total_denomintation_contribs.csv")

write_csv(clergy_wide, "clergy_wide_data.csv")

clergy_total <- clergy_wide |> 
  filter(!is.na(denomination)) |> distinct(contributor_name, denomination, affiliation)

```
```{r}
clergy_total |> 
  group_by(denomination) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))
  
  clergy_affilliation <- clergy_total |> 
    group_by(denomination, affiliation) |> 
    summarise(
      count = n()
    ) |> 
  rename(total_amount = count) |> 
  pivot_wider(names_from = affiliation, values_from = total_amount) |> 
    mutate(total_clergy = (DEMOCRAT + REPUBLICAN + `NA`)) |> 
    mutate(pct_democrat = (DEMOCRAT / total_clergy) * 100) |> 
    mutate(pct_republican = (REPUBLICAN / total_clergy) * 100)
  
total_contribution <- clergy_wide |> 
  mutate(total_contribution = (`2024`+ `2022` +`2020` + `2018` +  `2016`)) |> 
  filter(!is.na(denomination)) |> distinct(contributor_name, denomination, affiliation, total_contribution, zip5) 

total_contribution |> 
  filter(str_detect(affiliation, "DEMOCRAT")) |> 
  group_by(affiliation) |> 
  summarise(
    sum(total_contribution)
  )

total_contribution |> 
  filter(str_detect(affiliation, "REPUBLICAN")) |> 
  group_by(affiliation) |> 
  summarise(
    sum(total_contribution)
  )

clergy_affilliation_and_contributions <- total_contribution |> 
    group_by(denomination, total_contribution) |> 
    summarise(
      count = n()
    ) |> group_by(denomination) |> 
    summarise(
      sum(total_contribution)
    )

dem_clergy_affilliation_and_contributions <- total_contribution |>
  filter(str_detect(affiliation, "DEMOCRAT")) |>
  group_by(denomination) |>
  summarise(total = sum(total_contribution))

rep_clergy_affilliation_and_contributions <- total_contribution |>
  filter(str_detect(affiliation, "REPUBLICAN")) |>
  group_by(denomination) |>
  summarise(total = sum(total_contribution))

rep_clergy_affilliation_and_contributions |> 
  summarise(
    sum(total)
  )

dem_clergy_affilliation_and_contributions |> 
  summarise(
    sum(total)
  )
  

#I think all this code here goes pretty far (as far as I think I want to go) into answering question 1. There are a lot of cool stats (like retired clergy being the number one contributors by democratic clergy) in here and I think some good opportunities for visuals. I think that what I did here, potentially opened up questions which I hadn't thought to ask before about retired and self employed clergy contributions compared to clergy that are still working or work for a established denomination. The sheer amount that republicans are outspent too is also I think newsworthy if not unexpected. I also think that the stats about which denominations are mostly Democratic or Republican are newsworthy, and interesting, especially the fact that the Catholic Church is the most evenly split denomination in the state.#
```

```{r}
md_county_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              year = 2021,
              state = "MD",
              geometry = TRUE)

all_zipcodes <- get_acs(
  geography = "zcta",
  variables = "B01003_001",
  geometry = TRUE,
  year = 2021
)
#Got this with the help of a LLM#

md_zipcodes <- st_join(all_zipcodes, md_county_population, join = st_intersects) |> 
  select(GEOID.x, NAME.x, NAME.y) |> 
  rename(zip_code = GEOID.x, zip_name = NAME.x, county_name = NAME.y) |> 
  filter(str_detect(county_name, "Maryland"))

md_clergy_county <- total_contribution |> 
  inner_join(md_zipcodes, join_by(zip5 == zip_code)) |> 
  select(contributor_name, denomination, affiliation, total_contribution, county_name, zip5)

#This is some work I did to start to get at the geography question. I think that I've managed to get every clergy that I want assigned a county. Now I don't really understand why the code doubles the amount of clergy in the dataframe, but I think that something like this is at least going down the right track. This is the question I spent the second most time on, but I still have work to do here.#

```


```{r}
md_state_clergy_contribs <- read_csv("md_contribs.csv")

md_state_clergy_contribs <- md_state_clergy_contribs |> 
  mutate(committee_name = str_to_upper(committee_name)) |> 
  mutate(contributor_county_of_residence = str_to_upper(contributor_county_of_residence)) |> 
  mutate(committee_type = str_to_upper(committee_type)) 

#I at least wanted to load and start cleaning the data you provided last class. I'll keep working with this more for my last question#
```
Questions:
What denominations have the most clergy contributing money to political campaigns?

Are there certain cities or areas that produce clergy that contribute to political campaigns more than others?

How many of the clergy on the list are actually employed by or working for a real church? 

Do the outlier individual contributors have anything in common?

What are the general trends in giving by cycle? Are contributions increasing year to year, and what could that trend look like in the future? 

Do these clergy donors also give to Maryland state campaigns? 





