---
title: "data_analysis"
output: html_document
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tidycensus)
library(tigris)
library(sf)
library(ggthemes)
library(stringr)
library(scales)
```


```{r}
clergy_contribs_2016 <- read_csv("clergy_contribs_2016.csv")
clergy_contribs_2018 <- read_csv("clergy_contribs_2018.csv")
clergy_contribs_2020 <- read_csv("clergy_contribs_2020.csv")
clergy_contribs_2022 <- read_csv("clergy_contribs_2022.csv")
clergy_contribs_2024 <- read_csv("clergy_contribs_2024.csv")

clergy_contribs_combined <- bind_rows(list(clergy_contribs_2016, 
                                           clergy_contribs_2018, 
                                           clergy_contribs_2020, 
                                           clergy_contribs_2022, 
                                           clergy_contribs_2024)) |> 
  mutate(zip5 = str_sub(contributor_zip, start=1L, end=5L))

clergy_totals_by_cycle <- clergy_contribs_combined |>
  group_by(contributor_name, two_year_transaction_period, contributor_employer, zip5, committee_name) |>
  summarize(
    total_amount = sum(contribution_receipt_amount),
    num_contributions = n(),
  )

clergy_wide <- clergy_totals_by_cycle |>
  select(contributor_name, contributor_employer, two_year_transaction_period, total_amount, zip5, committee_name) |>
  pivot_wider(
    names_from = two_year_transaction_period,
    values_from = total_amount,
    values_fill = 0) |> 
  rename(denomination = contributor_employer) |> 
  mutate(
    denomination = case_when(
        str_detect(denomination, "BAPTIST") ~ "Baptist",
        str_detect(denomination, "CATHOLIC") ~ "Catholic",
        str_detect(denomination, "METHODIST") ~ "Methodist",
        str_detect(denomination, "PRESBY") ~ "Presbyterian",
        str_detect(denomination, "LUTHERAN") ~ "Lutheran",
        str_detect(denomination, "EPISCOPAL") ~ "Episcopal",
        str_detect(denomination, "AME") ~ "AME",
        str_detect(denomination, "ORTHODOX") ~ "Orthodox",
        str_detect(denomination, "CHURCH OF CHRIST") ~ "Church of Christ",
        str_detect(denomination, "ADVENTIST") ~ "Seventh-Day Adventist",
        str_detect(denomination, "SYNAGOGUE") ~ "Jewish",
        str_detect(denomination, "BUDDHIST") ~ "Buddhist",
        str_detect(denomination, "HINDU") ~ "Hindu",
        str_detect(denomination, "SIKH") ~ "Sikh",
        str_detect(denomination, "RETIRED") ~ "Retired",
        str_detect(denomination, "SELF") ~ "Self Employed",
        str_detect(denomination, "ARCHDIOCESE") ~ "Catholic",
        str_detect(denomination, "CONFERENCE") ~ "Seventh-Day Adventist",
        .default = NA
      )
  ) |> 
  mutate(
    affiliation = case_when(
      str_detect(committee_name, "ACTBLUE") ~ "DEMOCRAT",
      str_detect(committee_name, "HILLARY") ~ "DEMOCRAT",
      str_detect(committee_name, "KAMALA") ~ "DEMOCRAT",
      str_detect(committee_name, "BIDEN") ~ "DEMOCRAT",
      str_detect(committee_name, "FETTERMAN") ~ "DEMOCRAT",
      str_detect(committee_name, "DSCC") ~ "DEMOCRAT",
      str_detect(committee_name, "DEMOCRATIC") ~ "DEMOCRAT",
      str_detect(committee_name, "STOP REPUBLICAN") ~ "DEMOCRAT",
      str_detect(committee_name, "STOP") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "DCCC") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "ALSOBROOKS") ~ "DEMOCRAT",
      str_detect(committee_name, "BERNIE") ~ "DEMOCRAT",
      str_detect(committee_name, "RASKINS") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "REPUBLICAN & CONVENTION") ~ "REPUBLICAN",
      str_detect(committee_name, "MAKE AMERICA GREAT AGAIN") ~ "REPUBLICAN",
      str_detect(committee_name, "TRUMP") ~ "REPUBLICAN",
      str_detect(committee_name, "WINRED") ~ "REPUBLICAN",
      str_detect(committee_name, "RAND") ~ "REPUBLICAN",
      str_detect(committee_name, "REPUBLICAN NATIONAL COMMITTEE") ~ "REPUBLICAN",
      str_detect(committee_name, "NEVER SURRENDER") ~ "REPUBLICAN",
      str_detect(committee_name, "THE LINCOLN PROJECT") ~ "REPUBLICAN",
      .default = NA
    )
)
  

total_denomination_contribs <- clergy_wide |> 
  group_by(denomination) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))

total_party_affiliation <- clergy_wide |> 
  group_by(affiliation) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))

write_csv(total_denomination_contribs, "total_denomintation_contribs.csv")

write_csv(clergy_wide, "clergy_wide_data.csv")

clergy_total <- clergy_wide |> 
  filter(!is.na(denomination)) |> distinct(contributor_name, denomination, affiliation)

```

Question 1: What denominations have the most clergy contributing money to political campaigns?
```{r}
clergy_total |> 
  group_by(denomination) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))
  
  clergy_affilliation <- clergy_total |> 
    group_by(denomination, affiliation) |> 
    summarise(
      count = n()
    ) |> 
  rename(total_amount = count) |> 
  pivot_wider(names_from = affiliation, values_from = total_amount) |> 
    mutate(total_clergy = (DEMOCRAT + REPUBLICAN + `NA`)) |> 
    mutate(pct_democrat = (DEMOCRAT / total_clergy) * 100) |> 
    mutate(pct_republican = (REPUBLICAN / total_clergy) * 100)
  
total_contribution <- clergy_wide |> 
  mutate(total_contribution = (`2024`+ `2022` +`2020` + `2018` +  `2016`)) |> 
  filter(!is.na(denomination)) |> 
  group_by(contributor_name, denomination, affiliation, zip5) |> 
  summarize(grand_total = sum(total_contribution))

  #distinct(contributor_name, denomination, affiliation, total_contribution, zip5) 

total_contribution |> 
  filter(str_detect(affiliation, "DEMOCRAT")) |> 
  group_by(affiliation) |> 
  summarise(
    sum(grand_total)
  )

total_contribution |> 
  filter(str_detect(affiliation, "REPUBLICAN")) |> 
  group_by(affiliation) |> 
  summarise(
    sum(grand_total)
  )

clergy_affilliation_and_contributions <- total_contribution |> 
    group_by(denomination, grand_total) |> 
    summarise(
      count = n()
    ) |> group_by(denomination) |> 
    summarise(
      sum(grand_total)
    )

dem_clergy_affilliation_and_contributions <- total_contribution |>
  filter(str_detect(affiliation, "DEMOCRAT")) |>
  group_by(denomination) |>
  summarise(total = sum(grand_total))

rep_clergy_affilliation_and_contributions <- total_contribution |>
  filter(str_detect(affiliation, "REPUBLICAN")) |>
  group_by(denomination) |>
  summarise(total = sum(grand_total))

rep_clergy_affilliation_and_contributions |> 
  summarise(
    sum(total)
  )

dem_clergy_affilliation_and_contributions |> 
  summarise(
    sum(total)
  )

write_csv(total_contribution, "total_contribution.csv")

```
Question 2: Are there certain cities or areas that produce clergy that contribute to political campaigns more than others?
```{r}
unique_total_contribution <- total_contribution |> 
  mutate(contributor_name = str_to_upper(str_trim(contributor_name))) |> 
  mutate(clean_name = str_remove(contributor_name, " [A-Z]$")) |> 
  group_by(clean_name, denomination) |> 
  summarize(
    total_contributions = sum(grand_total, na.rm = TRUE),
    num_donations = n(),
    affiliation = first(affiliation) 
  ) |> 
  arrange(desc(total_contributions))

unique_contributions_zips <- unique_total_contribution |> 
  inner_join(total_contribution, join_by(clean_name == contributor_name)) |> 
  distinct(clean_name, zip5)

md_county_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              year = 2021,
              state = "MD",
              geometry = TRUE)

all_zipcodes <- get_acs(
  geography = "zcta",
  variables = "B01003_001",
  geometry = TRUE,
  year = 2021
)
#Got this with the help of a LLM#

md_zipcodes <- st_join(all_zipcodes, md_county_population, join = st_intersects) |> 
  select(GEOID.x, NAME.x, NAME.y) |> 
  rename(zip_code = GEOID.x, zip_name = NAME.x, county_name = NAME.y) |> 
  filter(str_detect(county_name, "Maryland"))

md_clergy_county <- unique_contributions_zips |> 
  inner_join(md_zipcodes, join_by(zip5 == zip_code)) |> 
  select(clean_name, county_name, zip5, geometry) |> 
  inner_join(unique_total_contribution, join_by(clean_name)) |> 
  select(clean_name, total_contributions, num_donations, affiliation, county_name, zip5, geometry,) |> 
  distinct(clean_name, .keep_all = TRUE)

md_county_contribs <- md_clergy_county |> 
  group_by(county_name) |> 
  summarise(
    total_county_contribs = sum(total_contributions),
    total_donations = sum(num_donations)
  ) |> 
  arrange(desc(total_county_contribs)) |> 
  mutate(clean_county = str_remove(county_name, "County, Maryland"),
         clean_county = str_remove(clean_county, "city, Maryland"),
         clean_county = str_trim(clean_county))

md_counties <- counties() |> 
  filter(STATEFP == "24")

county_contrib_map <- md_counties |> 
  left_join(md_county_contribs, join_by(NAME == clean_county)) |>
  select(NAMELSAD, GEOID, total_county_contribs, total_donations, geometry) |> 
  rename(full_county_name = NAMELSAD) |> 
  arrange(desc(total_county_contribs)) |>
  filter(!(full_county_name == "Baltimore city" & total_county_contribs > 100000)) |> 
  filter(!(full_county_name == "Baltimore County" & total_county_contribs < 7000))
#This is something I figured out with an llm, I messed up my joining when it came to Baltimore versus Baltimore City, and had to filter duplicates out.#

write_csv(md_clergy_county, "md_clergy_county.csv")
write_csv(county_contrib_map, "county_contrib_map.csv")

```
Total Contributions Map:
https://datawrapper.dwcdn.net/MJD0d/1/

Total Donations Map:
https://datawrapper.dwcdn.net/tX6IK/2/


Question 3: How many of the clergy on the list are actually employed by or working for a real church? (Or are not).
```{r}
total_clergy_by_denom <- unique_total_contribution |> 
  group_by(denomination) |> 
  summarise(
    num_clergy = n()
  ) |> 
  arrange(desc(num_clergy))

total_clergy_by_denom

total_clergy_contribution_by_denom <- unique_total_contribution |> 
  group_by(denomination) |> 
  summarise(
    num_clergy = n(),
    total_denom_contrib = sum(total_contributions)
  ) |> 
  arrange(desc(total_denom_contrib))

total_clergy_contribution_by_denom

total_clergy_by_denom |> 
  summarise(
    total_clergy = sum(num_clergy), 
    pct_retired_or_se = ((182 + 56) / (sum(num_clergy)) * 100)
  )

total_clergy_contribution_by_denom |> 
  summarise(
    total_money_contrib = sum(total_denom_contrib),
    amt_contributed_by_retired_or_se = (74467.83 + 60815.54), 
    pct_contributed_by_retired_or_se = ((74467.83 + 60815.54) / (sum(total_denom_contrib)) *100)
  )

write_csv(total_clergy_by_denom, "total_clergy_by_denom.csv")

write_csv(total_clergy_contribution_by_denom, "total_clergy_contribution_by_denom.csv")
```
Self Employed and Retired:
https://datawrapper.dwcdn.net/HUqYX/1/ 

Self Employed and Retired Contributions:
https://datawrapper.dwcdn.net/1psMG/1/ 


Question 4: Do the outlier individual contributors have anything in common?
```{r}
biggest_clergy_contribs <- unique_total_contribution |> 
  head(100)

most_denoms_donations_in_top100 <- biggest_clergy_contribs |> 
  group_by(denomination) |> 
  summarise(
    num_donations = n()
  ) |> 
  arrange(desc(num_donations))

most_denoms_donations_in_top100

most_money_contributed_in_top100_by_denom <- biggest_clergy_contribs |> 
  group_by(denomination) |> 
  summarise(
    amt_contributed = sum(total_contributions)
  ) |> 
  arrange(desc(amt_contributed))

most_money_contributed_in_top100_by_denom 

most_money_SANS_DORIS <- biggest_clergy_contribs |> 
  filter(!str_detect(clean_name, "COWAN")) |> 
  group_by(denomination) |> 
  summarise(
    amt_contributed = sum(total_contributions)
  ) |>
  arrange(desc(amt_contributed))
  
most_money_SANS_DORIS

most_money_contributed_in_top100_by_denom  |>
  ggplot() +
  geom_bar(aes(x=reorder(denomination,  amt_contributed), weight= amt_contributed)) +
  coord_flip() +
  theme_economist() +
  labs(
    title=str_wrap("Amount of money donated by denominations based on the top 100 clergy donators", width = 40),
    x = "Denominations",
    y = "Money Contributed",
    caption = "Federal Election Commission"
  )

most_money_SANS_DORIS  |>
  ggplot() +
  geom_bar(aes(x=reorder(denomination,  amt_contributed), weight= amt_contributed)) +
  coord_flip() +
  theme_economist() +
  labs(
    title=str_wrap("Amount of money donated by denominations based on the top 100 clergy donators, with Doris Cownan     removed", width = 40),
    x = "Denominations",
    y = "Money Contributed",
    caption = "Federal Election Commission"
  )
  
```

Question 5: What are the general trends in giving by cycle? Are contributions increasing year to year, and what could that trend look like in the future? 
```{r}
donation_amt_per_cycle <- clergy_wide |> 
  pivot_longer(
    cols = c(`2016`, `2018`, `2020`, `2022`, `2024`),
    names_to = "election_cycle",
    values_to = "amt_donated"
  ) |> 
  group_by(election_cycle) |> 
  summarise(
    total_amt_donated = sum(amt_donated)
  )

donation_amt_per_cycle

donation_amt_per_cycle |>
  mutate(election_cycle = as.numeric(election_cycle)) |> 
  ggplot() +
  geom_line(aes(x=election_cycle, y=total_amt_donated)) +
  geom_point(aes(x=election_cycle, y=total_amt_donated)) +
  scale_y_continuous(labels = label_comma()) +
  theme_minimal() +
  labs(
    title="Clergy contributions per election cycle generally rising",
    x = "Election Cycle",
    y = "Total Amount Donated",
    caption = "Federal Election Commission"
  )

non_election_year_contribution <- donation_amt_per_cycle |> 
  filter(!str_detect(election_cycle, "2016|2020|2024")) |> 
    pivot_wider(
    names_from = election_cycle,
    values_from = total_amt_donated,
    values_fill = 0
  )

election_year_contribution <- donation_amt_per_cycle |> 
  filter(!str_detect(election_cycle, "2018|2022")) |> 
    pivot_wider(
    names_from = election_cycle,
    values_from = total_amt_donated,
    values_fill = 0
  )

election_year_contribution |> 
  summarise(
    pct_change_2016_2020 = ((`2020` - `2016`)/`2016`)*100,
    pct_change_2020_2024 = ((`2024` - `2020`)/`2020`)*100,
    pct_change_2016_2024 = ((`2024` - `2016`)/`2016`)*100
  )

non_election_year_contribution |> 
  summarise(
    pct_change_2018_2022 = ((`2022` - `2018`)/`2018`)*100
  )
```



```{r}
md_state_clergy_contribs <- read_csv("md_contribs.csv")

md_state_clergy_contribs <- md_state_clergy_contribs |> 
  mutate(committee_name = str_to_upper(committee_name)) |> 
  mutate(contributor_county_of_residence = str_to_upper(contributor_county_of_residence)) |> 
  mutate(committee_type = str_to_upper(committee_type)) 
```


Questions:
What denominations have the most clergy contributing money to political campaigns?

Are there certain cities or areas that produce clergy that contribute to political campaigns more than others?

How many of the clergy on the list are actually employed by or working for a real church? 

Do the outlier individual contributors have anything in common?

What are the general trends in giving by cycle? Are contributions increasing year to year, and what could that trend look like in the future? 

Do these clergy donors also give to Maryland state campaigns? 





