---
title: "data_analysis"
output: html_document
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tidycensus)
library(tigris)
```


```{r}
clergy_contribs_2016 <- read_csv("clergy_contribs_2016.csv")
clergy_contribs_2018 <- read_csv("clergy_contribs_2018.csv")
clergy_contribs_2020 <- read_csv("clergy_contribs_2020.csv")
clergy_contribs_2022 <- read_csv("clergy_contribs_2022.csv")
clergy_contribs_2024 <- read_csv("clergy_contribs_2024.csv")

clergy_contribs_combined <- bind_rows(list(clergy_contribs_2016, 
                                           clergy_contribs_2018, 
                                           clergy_contribs_2020, 
                                           clergy_contribs_2022, 
                                           clergy_contribs_2024)) |> 
  mutate(zip5 = str_sub(contributor_zip, start=1L, end=5L))

clergy_totals_by_cycle <- clergy_contribs_combined |>
  group_by(contributor_name, two_year_transaction_period, contributor_employer, zip5, committee_name) |>
  summarize(
    total_amount = sum(contribution_receipt_amount),
    num_contributions = n(),
  )

clergy_wide <- clergy_totals_by_cycle |>
  select(contributor_name, contributor_employer, two_year_transaction_period, total_amount, zip5, committee_name) |>
  pivot_wider(
    names_from = two_year_transaction_period,
    values_from = total_amount,
    values_fill = 0) |> 
  rename(denomination = contributor_employer) |> 
  mutate(
    denomination = case_when(
        str_detect(denomination, "BAPTIST") ~ "Baptist",
        str_detect(denomination, "CATHOLIC") ~ "Catholic",
        str_detect(denomination, "METHODIST") ~ "Methodist",
        str_detect(denomination, "PRESBY") ~ "Presbyterian",
        str_detect(denomination, "LUTHERAN") ~ "Lutheran",
        str_detect(denomination, "EPISCOPAL") ~ "Episcopal",
        str_detect(denomination, "AME") ~ "AME",
        str_detect(denomination, "ORTHODOX") ~ "Orthodox",
        str_detect(denomination, "CHURCH OF CHRIST") ~ "Church of Christ",
        str_detect(denomination, "ADVENTIST") ~ "Seventh-Day Adventist",
        str_detect(denomination, "SYNAGOGUE") ~ "Jewish",
        str_detect(denomination, "BUDDHIST") ~ "Buddhist",
        str_detect(denomination, "HINDU") ~ "Hindu",
        str_detect(denomination, "SIKH") ~ "Sikh",
        str_detect(denomination, "RETIRED") ~ "Retired",
        str_detect(denomination, "SELF") ~ "Self Employed",
        str_detect(denomination, "ARCHDIOCESE") ~ "Catholic",
        str_detect(denomination, "CONFERENCE") ~ "Seventh-Day Adventist",
        .default = NA
      )
  ) |> 
  mutate(
    affiliation = case_when(
      str_detect(committee_name, "ACTBLUE") ~ "DEMOCRAT",
      str_detect(committee_name, "HILLARY") ~ "DEMOCRAT",
      str_detect(committee_name, "KAMALA") ~ "DEMOCRAT",
      str_detect(committee_name, "BIDEN") ~ "DEMOCRAT",
      str_detect(committee_name, "FETTERMAN") ~ "DEMOCRAT",
      str_detect(committee_name, "DSCC") ~ "DEMOCRAT",
      str_detect(committee_name, "DEMOCRATIC") ~ "DEMOCRAT",
      str_detect(committee_name, "STOP REPUBLICAN") ~ "DEMOCRAT",
      str_detect(committee_name, "STOP") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "DCCC") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "ALSOBROOKS") ~ "DEMOCRAT",
      str_detect(committee_name, "BERNIE") ~ "DEMOCRAT",
      str_detect(committee_name, "RASKINS") ~ "DEMOCRAT",
      str_detect(committee_name, "HARRIS") ~ "DEMOCRAT",
      str_detect(committee_name, "REPUBLICAN & CONVENTION") ~ "REPUBLICAN",
      str_detect(committee_name, "MAKE AMERICA GREAT AGAIN") ~ "REPUBLICAN",
      str_detect(committee_name, "TRUMP") ~ "REPUBLICAN",
      str_detect(committee_name, "WINRED") ~ "REPUBLICAN",
      str_detect(committee_name, "RAND") ~ "REPUBLICAN",
      str_detect(committee_name, "REPUBLICAN NATIONAL COMMITTEE") ~ "REPUBLICAN",
      str_detect(committee_name, "NEVER SURRENDER") ~ "REPUBLICAN",
      .default = NA
    )
)
  

total_denomination_contribs <- clergy_wide |> 
  group_by(denomination) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))

#This code I think may help me on questions 1 and 3. 1 because I want to work on seeing the majoirty party affiliation for individual denominations in Maryland, and 3 because it's at least a start at trying to county the retired or self employed clergy. The problem is all the NAs and I'm kind of lost on how to lower that, I think case_when might be limited here.#

total_party_affiliation <- clergy_wide |> 
  group_by(affiliation) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))

#This is also code aimed toward question 1 and trying to determine party affiliation.#

denomination_vs_affiliation <- 

write_csv(total_denomination_contribs, "total_denomintation_contribs.csv")

write_csv(clergy_wide, "clergy_wide_data.csv")

clergy_total <- clergy_wide |> 
  filter(!is.na(denomination)) |> distinct(contributor_name, denomination, affiliation)

```
```{r}
clery_total |> 
  group_by(denomination) |> 
  summarise(
    count = n()
  ) |> 
  rename(total_amount = count) |> 
  arrange(desc(total_amount))
  
  clergy_affilliation <- clergy_total |> 
    group_by(denomination, affiliation) |> 
    summarise(
      count = n()
    ) |> 
  rename(total_amount = count) |> 
  pivot_wider(names_from = affiliation, values_from = total_amount) |> 
    mutate(total_clergy = (DEMOCRAT + REPUBLICAN + `NA`)) |> 
    mutate(pct_democrat = (DEMOCRAT / total_clergy) * 100) |> 
    mutate(pct_republican = (REPUBLICAN / total_clergy) * 100)
  
```

```{r}
md_state_clergy_contribs <- read_csv("md_contribs.csv")

md_state_clergy_contribs <- md_state_clergy_contribs |> 
  mutate(committee_name = str_to_upper(committee_name)) |> 
  mutate(contributor_county_of_residence = str_to_upper(contributor_county_of_residence)) |> 
  mutate(committee_type = str_to_upper(committee_type)) 

#I at least wanted to load and start cleaning the data you provided last class. I'll keep working with this more for my last question#
```
Questions:
What denominations have the most clergy contributing money to political campaigns?

Are there certain cities or areas that produce clergy that contribute to political campaigns more than others?

How many of the clergy on the list are actually employed by or working for a real church? 

Do the outlier individual contributors have anything in common?

What are the general trends in giving by cycle? Are contributions increasing year to year, and what could that trend look like in the future? 

Do these clergy donors also give to Maryland state campaigns? 





