---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
rls_streaming <- read_csv("data/rls_streaming.csv", col_names = FALSE, skip = 1) |> 
  rename(
    weight = X1, 
    in_person_attendence = X2,
    streaming_attendence = X3,
    education_level = X4,
    household_income = X5, 
    ethnicity = X6,
    race = X7,
    gender = X8,
    birthdecade = X9
  )


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Education does not meaningfully distinguish Americans in terms of weekly in-person religious attendance, but it sharply differentiates who participates online.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
rls_streaming |> 
  filter(streaming_attendence == 2 ) |> 
  filter(education_level == 1 | education_level == 2)

rls_streaming |> 
  filter(streaming_attendence == 2) |> 
  filter(education_level == 3 | education_level == 4)

rls_streaming |> 
  filter(in_person_attendence == 2) |> 
  filter(education_level == 1 | education_level == 2)

rls_streaming |> 
  filter(in_person_attendence == 2) |> 
  filter(education_level == 3 | education_level == 4)

# Display results of code below this codeblock


```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-