---
title: "lab_07"
author: "Derek Willis"
date: "2025-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(tidycensus)
library(janitor)

```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key only if you haven't already. You won't need to install or overwrite it.

```{r echo=FALSE}
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2021 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2021 5-year ACS (which is called "acs5") and save them to a dataframe

```{r}
acs5 <- load_variables(2021, "acs5", cache = TRUE)
```

## Answer questions

**Q1** What is the Maryland county with the highest percentage of the population with a bachelor's degree or higher, according to the ACS 2021 5-year estimates? The variable you want is `DP02_0068P` from the ACS Data Profile tables — it's the pre-calculated percentage of the population 25 and over with a bachelor's degree or higher.

The county with the highest percentage is Howard County. 

**A1**

```{r}
maryland_degrees <- get_acs(
  geography = "county",
  variables = "DP02_0068P",
  year = 2021) |> 
  filter(str_detect(NAME, "Maryland")) |> 
  arrange(desc(estimate))

maryland_degrees

```

**Q2** Now let's join that data with U.S. Religion Census data. Load the `usreligioncensus_county_2020.csv` file and filter it for Maryland counties. Then join it to your education data. Look at the top 5 counties by educational attainment and the bottom 5. How do their religious adherence rates compare? Write a sentence describing what you see.

**A2**

Comparing the top 5 counties based on educational attainment I think that you could generally say that the higher educated the county is, generally it will be more religiously adherent than it's bottom five counterpart. However, the two are not really that far apart.

```{r}
md_religous_counties <- read_csv("data/usreligioncensus_county_2020.csv") |> clean_names() |> 
  filter(str_detect(state_name, "Maryland"))

md_religion_education <- maryland_degrees |> 
 inner_join(md_religous_counties, join_by(GEOID == fips)) |> 
  select(county_name, estimate, adherents_as_percent_of_population) |> 
  arrange(desc(estimate))

md_religion_education |> 
  head(5)

md_religion_education |> 
  tail(5)

md_religion_education
             
```

**Q3** Using your joined Maryland county data, find the 5 most and 5 least religiously diverse counties based on the Diversity Index Value. Show the largest religious body by number of adherents and the median household income (you'll need to get income from the ACS and join it in) for those counties. What patterns do you notice about which religious groups dominate in more vs. less diverse counties?

The Catholic church dominates in the more diverse counties (they also kind of dominate in the less diverse as well, but not as much). I would imagine this has to do with the fact that the Catholic church has massive inroads in the Hispanic communities, and those Hispanic communities are what make those counties more diverse.

**A3**

```{r}
md_religion_education <- maryland_degrees |> 
 inner_join(md_religous_counties, join_by(GEOID == fips)) 

income_data <-  get_acs(
  geography = "county",
  variables = c(medincome = "B19013_001"),
  state = "MD",
  year = 2021) |> 
  rename(median_income = estimate)

md_religion_income_combined <- md_religion_education |> 
  inner_join(income_data, join_by(GEOID)) |> 
  select(county_name, diversity_index_value, largest_body_by_number_of_adherents, median_income, number_of_adherents) 

md_religion_income_combined |> 
  arrange(desc(diversity_index_value)) |> 
  select(county_name, largest_body_by_number_of_adherents, diversity_index_value, median_income) |> 
  head(5)

md_religion_income_combined |> 
  arrange(desc(diversity_index_value)) |> 
  select(county_name, largest_body_by_number_of_adherents, diversity_index_value, median_income) |> 
  tail(5)


```

**Q4** Is there a relationship between wealth and the number of religious congregations in a community? Using your joined data that includes median household income, look at the "Congregations Per 100,000 Population" column. Arrange the counties by congregations per capita. Do the counties with the most congregations per capita tend to be wealthier or less wealthy? What about the counties with the fewest? Write a sentence describing the pattern.

**A4**

Countries with more congregations per 100,000 actually tend to be less wealthy. The wealthier counties like Howard, and Montgomery County have signifigantly less congregations per 100,000 than the counties that are far less wealthy and more rural like Garrett County.

```{r}
md_religion_income_gdp <- md_religion_education |> 
  inner_join(income_data, join_by(GEOID)) |> 
  select(county_name, congregations_per_100_000_population, median_income) |> 
  arrange(desc(congregations_per_100_000_population))

md_religion_income_gdp |> 
  head(5)

md_religion_income_gdp |> 
  tail(5)
```

**Q5** Let's compare religion and income across all states. Get state-level median household income from the ACS and join it with the state-level religion census data (`usreligioncensus_state_2020.csv`). Remember you'll need to make the FIPS codes match in format, as we did in the pre_lab.

Arrange the states by adherence rate. Among the 10 states with the highest adherence rates, what is the average median household income? Among the 10 states with the lowest adherence rates, what is the average? What regional patterns do you notice in each group?

**A5**

In the highest adherence states the average median income is 62914.1. In the lowest the average is 68750.8. The adherence rate seems to be much higher in Southern states, and lower in Midwestern and Northern states, as well as some East/Eastcoast states. 

```{r}
state_religion_data <- read_csv("data/usreligioncensus_state_2020.csv") |> clean_names()

state_median_income <- get_acs(
  geography = "state",
  variables = "B19013_001",
  year = 2021) |> 
  rename(median_income = estimate)

state_data_combined <- state_religion_data |> 
  inner_join(state_median_income, join_by(state_code == GEOID)) |> 
  select(state_name, adherents_as_percent_of_population, median_income) |> 
  arrange(desc(adherents_as_percent_of_population))

state_data_combined |> 
  head(10)

state_data_combined |> 
  head(10) |> 
  summarise(mean(median_income))

state_data_combined |> 
  tail(10)

state_data_combined |> 
  tail(10) |> 
  summarise(mean(median_income))
```

**Q6** Your own question: Come up with your own question that combines Census API data with U.S. Religion Census data. Your question should:

-   Use at least one ACS variable we haven't used in previous questions
-   Involve a join between Census and religion data
-   Include grouping, ranking, or comparison across geographic areas

Write your question, show your code, and provide an answer with interpretation.

**A6**

Your question:

On a statewide level which congregations are the wealthiest. To do this I will arrange according median household income, while comparing to largest body by number of adherents, looking at the top ten lowest and the top ten highest.

```{r}
state_religion_data <- read_csv("data/usreligioncensus_state_2020.csv") |> clean_names()

state_median_income <- get_acs(
  geography = "state",
  variables = "B19013_001",
  year = 2021) |> 
  rename(median_income = estimate)

state_data_combined <- state_religion_data |> 
  inner_join(state_median_income, join_by(state_code == GEOID)) |> 
  select(state_name, largest_body_by_number_of_adherents, median_income) |> 
  arrange(desc(median_income))

state_data_combined |> 
  head(10)

state_data_combined |> 
  tail(10)

```

Your answer and interpretation:

Looking at the top ten wealthiest states, the dominant church by number of adherents is the Catholic church by far with only the Washington state being in the top ten. In comparison the bottom median income states are all primarily Southern Baptist with the exception of New Mexico, Louisiana, and West Virginia.

Now I don't think that this is exactly indicative of church wealth. However, I do think that it shows that in the states that have higher median incomes, that the biggest faith is Catholicism, which is interesting. The richer states at least according to this seems to be more eastern and northern (with the exception of Hawaii and Cali), and Catholic, while the poorer states are Southern and protestant, primarily baptist. This I also think is intersting to think about politically, where the thing that links the majority of the top ten together is that they are all heavily democratic states, and what links the majority of the bottom ten together is that they are heavily republican states. 
